---
title: "Energy_analysis_index_system"
author: "XinzeTang"
date: "2019/12/21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages
```{r warning=FALSE}
library(dplyr)
library(ggplot2)
```

## Load data which has been preprocessed. 

```{r echo=FALSE}
energy <- read.csv('wide_data_fill_final.csv', header = T)
columns <- names(energy)
energy
```


## Calculate indexes
### Environmental Sustainability
1. Renewable Energy Production Ratio
```{r}
## Extract production data
production <- energy %>%
  select(Year, StateCode, BMTCB, GETCB, HYTCB, WYTCB, SOTCB, NUETB, CLPRB, PAPRB, NGMPB, TEPRB) %>%
  mutate(Total_Production = BMTCB + GETCB + HYTCB + WYTCB + SOTCB + NUETB + CLPRB + PAPRB + NGMPB) %>%
  mutate(Renew_Ratio = (BMTCB + GETCB + HYTCB + WYTCB + SOTCB)/Total_Production)
```
```{r}
ggplot(data = production, mapping = aes(x = Year, y = Renew_Ratio))+
  geom_line(mapping = aes(color = StateCode)) +
  labs(y = "Renewable Energy Production Ratio")
```
2. Renewable Energy Consumption Ratio
```{r}
## Extract consumption data
consumption <- energy %>%
  select(Year, StateCode, BMTCB, CLTCB, GETCB, HYTCB, NNTCB, PMTCB, WYTCB, SOTCB, NUETB, TETCB) %>%
  mutate(Total_Consumption = BMTCB + CLTCB + GETCB + HYTCB + NNTCB + PMTCB + WYTCB + SOTCB + NUETB) %>%
  mutate(Renew_Ratio = (BMTCB + GETCB + HYTCB + WYTCB + SOTCB)/Total_Consumption)
```
```{r}
ggplot(data = consumption, mapping = aes(x = Year, y = Renew_Ratio))+
  geom_line(mapping = aes(color = StateCode)) +
  labs(y = "Renewable Energy Consumption Ratio")
```


3. Efficiency (Factor of Converting)
  + Because the energy department only provide factor of converting for three main energy: natural gas, petroleum and coal, so we can onl use these three factors. 
```{r}
## MinMaxScaler
rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}
energy <- energy %>%
  mutate(
    CLEIK_scaled = rescale01(CLEIK),
    NGEIK_scaled = rescale01(NGEIK),
    PAEIK_scaled = rescale01(PAEIK)
  )
```
```{r}
## Calculate efficiency
efficiency <- energy %>%
  select(Year, StateCode, CLEIK_scaled, NGEIK_scaled, PAEIK_scaled) %>%
  mutate(Efficiency = (CLEIK_scaled+NGEIK_scaled+PAEIK_scaled)/3)
efficiency
```
```{r}
ggplot(data = efficiency, mapping = aes(x = Year, y = Efficiency))+
  geom_line(mapping = aes(color = StateCode)) +
  labs(y = "Efficiency (Factor of Converting, scaled to 0-1)")
```

4. CO2 Emission per unit energy use
```{r echo=FALSE}
co2az = read.csv('co2_ar_fill_1.csv')
co2ca = read.csv('co2_ca_fill_1.csv')
co2nm = read.csv('co2_me_fill_1.csv')
co2tx = read.csv('co2_te_fill_1.csv')
co2az['StateCode']='AZ'
co2ca['StateCode']='CA'
co2nm['StateCode']='NM'
co2tx['StateCode']='TX'
```
```{r}
## Bind data from different files
co2_emission <- rbind(co2az, co2ca, co2nm, co2tx) %>%
  select(year, StateCode, Total) %>%
  rename(Year=year, CO2_Emission=Total) %>%
  filter(Year<=2009)
## Extract consumption data
consumption <- energy %>%
  select(Year, StateCode, BMTCB, CLTCB, GETCB, HYTCB, NNTCB, PMTCB, WYTCB, SOTCB, NUETB, TETCB) %>%
  mutate(Total_Consumption = BMTCB + CLTCB + GETCB + HYTCB + NNTCB + PMTCB + WYTCB + SOTCB + NUETB) %>%
  select(Year, StateCode, Total_Consumption, TETCB)
## Take the log of consumption to prevent small values
co2_emission <- co2_emission %>% 
  left_join(consumption, by = c('Year', 'StateCode')) %>%
  mutate(Unit_CO2_Emission = CO2_Emission/log(Total_Consumption))
  
## MinMaxScaler
co2_emission <- co2_emission %>%
  mutate(Unit_CO2_Emission_scaled = rescale01(Unit_CO2_Emission))
```
```{r}
ggplot(data = co2_emission, mapping = aes(x = Year, y = Unit_CO2_Emission_scaled))+
  geom_line(mapping = aes(color = StateCode)) +
  labs(y = "Unit CO2 Emission (scaled to 0-1)")
```

### Economic Growth and Development
1. Energy Intensity
2. Export Contribution
3. Import Contribution
4. Affordability

### Energy Access and Security
1. Self Sufficiency
2. Energy Diversity Index
3. Price Stability
4. Energy Production per capita

